---
tittle: "p8105 homework 2"
anthor: Yujin Zhang
output: github_document
---

This is a R MARKDOWN file for homework 2.

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```

## Problem 1 read and clean the Mr. Trash Wheel dataset.
```{r}
#read and clean Mr. Trash Wheel dataset
trashwheel_df = 
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", range = "A2:N574") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%  # omit rows that don't include dumpster data
  mutate(sports_balls = as.integer(sports_balls))  #round to the nearest integer

# read and clean the precipitation data
precipitation_2018_df =
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na(total) %>%  # omit rows that don't include data
  mutate(year = 2018) %>%  # add a 'year' variable
  relocate(year)

precipitation_2019_df = 
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na(total) %>%  # omit rows that don't include data
  mutate(year = 2019) %>%  # add a 'year' variable
  relocate(year)

# combine two datasets
precipitation_com_df = 
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>% 
  mutate(month = month.name[month])  # convert month to a character variable
```

```{r, echo=FALSE}
# for calculating the median number of sports balls
sportsballs_2019_df = filter(trashwheel_df, year == 2019)
```

The number of observation in trashwheel_df dataset is `r nrow(trashwheel_df)`, and the key variables are: `r names(trashwheel_df)`

The number of observation in precipitation_com_df dataset is `r nrow(precipitation_com_df)`. The key variables are: `r names(precipitation_com_df)`

The total precipitation in 2018 is `r sum(pull(precipitation_2018_df, total))`. 

The median number of sports balls in a dumpster in 2019 is `r median(pull(sportsballs_2019_df, sports_balls))`.



## Problem 2
```{r}
# read and clean the pols_month.csv data
pols_month_df = 
  read_csv("./hw_data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%  # break up the variable 'mon' 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%  # transfer the variables to integer
  mutate(month = month.name[month]) %>%  # use month name in table
  mutate(president = prez_dem) %>%  # create a 'president' variable
  mutate(president = recode(president, "0" = "gop", "1" = "dem"))  # take values for variable

pols_month_df = select(pols_month_df, -day, -prez_dem, -prez_gop)  # remove three variables
```


## Clean the data in snp.csv
```{r}
snp_df =
  read_csv("./hw_data/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(date = parse_datetime(date, "%m/%d/%y")) %>% # figure out date in another formula
  mutate(date = as.Date(ifelse(date > Sys.Date(), format(date, "19%y-%m-%d"), format(date)))) %>%  # correct wrong 'year' numbers
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  arrange(year, month) %>%  # arrange according to year and month
  mutate(month = month.name[month]) 

snp_df = select(snp_df, -day) %>%  # remove one variable
  relocate(year)

```

## Tide the unemployment data.
```{r}
unemployment_df = 
  read_csv("./hw_data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    January = jan,
    Febrary = feb,
    March = mar,
    April = apr,
    May = may,
    June = jun,
    July = jul,
    Augest = aug,
    September = sep,
    October = oct,
    November = nov,
    December = dec
  ) %>%  # rename variables
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "percentage of unemployment"
  )  # tidy the data to have the same variable with above two datasets
```

## Join the three dataset above.
```{r}
join_two_df = left_join(pols_month_df, snp_df, by = c("year" = "year", "month" = "month")) 
join_df = left_join(join_two_df, unemployment_df, by = c("year" = "year", "month" = "month"))
```

The dataset of "pols_month_df" contains 9 variables:`r names(pols_month_df)` and 822 observations with a range from 1947-2015.

The dataset of "snp_df" contains 3 variables:`r names(snp_df)` and 787 observations with a range from 1950-2015.

The dataset of "unemployment_df" contains 3 variables:`r names(unemployment_df)` and 816 observations with a range from 1948-2015.

The resulting dataset "join_df" contains variables `r names(join_df)` and 822 observations from 1947-2015.


## Problem 3
```{r}
popl_babynames_df =
  read_csv("./hw_data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>%
  mutate(childs_first_name = toupper(childs_first_name)) %>%  # unify the data format
  mutate(ethnicity = tolower(ethnicity)) %>% 
  mutate(gender = tolower(gender)) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander")) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic")) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic")) %>%
  distinct() # remove the duplicated rows

# create a table that shows the rank of the female name "Olivia"
olivia_popl_df =
  filter(popl_babynames_df, 
         gender == "female",
         childs_first_name == "OLIVIA"
         ) %>% 
  arrange(year_of_birth) %>% 
  rename(
    "year of birth" = year_of_birth,
    "childs first name" = childs_first_name
  )
knitr::kable(olivia_popl_df)

# create a table that shows the most popular name among male children over time
male_popl_df =
  filter(popl_babynames_df,
         gender == "male",
         rank == 1
         ) %>% 
  arrange(year_of_birth) %>% 
  rename(
    "year of birth" = year_of_birth,
    "childs first name" = childs_first_name
  )
knitr::kable(male_popl_df)
```

## Produce a scatter polt with male, non-hispanic children born in 2016
```{r}
malename_2016_df = 
  filter(popl_babynames_df,
         year_of_birth == 2016,
         gender == "male",
         ethnicity == "white non hispanic",
         )

ggplot(malename_2016_df, aes(x = rank, y = count)) + geom_point() + ggtitle("names' rank with numbers of male, white non-hispanic children born in 2016")
```
