---
tittle: "p8105 homework 2"
anthor: Yujin Zhang
output: github_document
---

This is a R MARKDOWN file.

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1 read and clean the Mr. Trash Wheel dataset.
```{r}
trashwheel_df = 
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster)
trashwheel_df = transform(trashwheel_df, sports_balls = as.integer(sports_balls)) #no idea

precipitation_2018_df =
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  rename(total_precipitation = total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precipitation_2019_df = 
  read_excel("./hw_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  rename(total_precipitation = total) %>% 
  mutate(year = 2019) %>% 
  relocate(year)

precipitation_df = 
  left_join(precipitation_2018_df, precipitation_2019_df, by = "month")
precipitation_df = transform(precipitation_df, month = month.name[month]) # no idea what else solution to combine the df
```

```{r, echo=FALSE}
sportsballs_2017_df = filter(trashwheel_df, year == 2017)
```

The number of observation in trash wheel dataset is`r nrow(trashwheel_df)`, and the number of observation in precipitation dataset is `r nrow(precipitation_2018_df) + nrow(precipitation_2019_df)`. The key variables?
The total precipitation in 2018 is `r sum(pull(precipitation_2018_df, total_precipitation))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(sportsballs_2017_df, sports_balls))`.



## Problem 2: